---
title             : "Comprehension of Sentential Negation in Toddlers"
shorttitle        : "Comprehension of Negation in Toddlers"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Previous research suggests that children understand truth-conditional negation 
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib","references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library(jpeg)
library(grid)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction

## Previous Research

<!--
DIMENTION 1 = NEGATION TYPE
No: Austin, Theakston, Lieven, & Tomasello (2014), Feiman, Mody, Sanborn, and Carey (2017); Nordmeyer and Frank (2014)
Not: Donaldson (1970), De Villier and Flusberg (1975), Pea (1980;1982), Kim (1985), Austin, Theakston, Lieven, & Tomasello (2014), Feiman, Mody, Sanborn, and Carey (2017), 
didn't: Reuter, Feiman, Snedeker (2018)
pas: de Carvalho, Barrault, and Christophe (2017)

DIMENTION 2 = ARGUMENT TYPE
Predicative NP + not: De Villier and Flusberg (1975), Pea (1980;1982), Kim (1985), de Carvalho, Barrault, and Christophe (2017)
Locative-PP + No/Not: Feiman, Mody, Sanborn, and Carey (2017), Austin, Theakston, Lieven, & Tomasello (2014)
Transitive-Verb + didn't: Reuter, Feiman, Snedeker (2018)
Det-No + Noun: Nordmeyer and Frank (2014)

DIMENTION 2 = METHOD
Picture verification + sentence-completion: Donaldson (1970), De Villier and Flusberg (1975)
Picture verification + judgments: Slobin (1966), Kim (1985)
Picture verification + spontaneous feedback: Pea (1980; 1982) 
Picture verificatioin + eye-tracking: Reuter, Feiman, Snedeker (2018), Nordmeyer and Frank (2014)
Hide-and-seek with search behavior: Feiman, Mody, Sanborn, and Carey (2017), Austin, Theakston, Lieven, & Tomasello (2014)
Picture verification with looking-time: de Carvalho, Barrault, and Christophe (2017)

Other: Donaldson (1970) - Hummer, Wimmer, & Antes (1993)
-->

Research on comprehension of negation has used multiple methods tested children's comprehension using differentiof negated identity (e.g. *this is not a banana*), negated location (**)

Slobin (1966) was the first to test children's comprehension of negation. He used a sentence verification task in which participants saw a picture, heard a positive or a negative sentence, and had to press one of two buttons to determine wehther the sentence was "right" or "wrong". 

Unfortunately Slobin did not report the details of data on children's errors. This is what Donaldson (1970) did and reported that five-year-olds make substantial errors. 

De Villier and Flusberg (1975) imrpoved on Donaldson's work and showed much earlier comprehension.

Pea (1982) and Kim (1985) were the first to talk about truth-functional negation and test it experimentally. 

Austin, Theakston, Lieven, & Tomasello (2014) pointed out that there may be pragmatic issues with denial negation in picture verification tasks. Instead they advocated for information regarding location of nonvisible objects. 

Feiman, Mody, Sanborn, and Carey (2017)

Reuter, Feiman, Snedeker (2018)

de Carvalho, Barrault, and Christophe (2017) tested comprehension of French negation in 18-month-olds using a word learning paradigm. 

The trend in experimental research on comprehension of negation in children has been one of improving methods and capturing earlier comprehension than before. 

## Current Study

This study tests toddler's comprehension of verbal negation in sentences that deny the identity of an entity (e.g. *this is not a cat*). 

# Experiment 1

In the first experiment we focused on todders' ability to react to the truth altering property of negation. We habituated toddlers to positive and truth-preserving adverbs like *really* as in "this is really a dog". Then we tested them with either the similar adverb *now* or the negative and truth-altering adverb *not*. We predicted that if toddlers are sensitive to the functional role of negation in altering the truth conditions of statemets, they should dishabituate when they notice the change in truth conditions that no longer match visual stimuli truthfully.

## Methods

We conducted a looking time study that paired linguistic audio stimuli with images of objects appearing on a screen. The linguistic stimuli were simple sentences of the general form "this is [adverb] a [noun]" (e.g. "this is really a ball" vs. "this is not a ball"). The images were objects described by the nouns (e.g. a ball). There were 6 adverbs (*really*,*indeed*,*only*,*just*,*now*, and *not*) and 3 nouns (*ball*, *dog*, and *shoe*) resulting in 18 total possible trials. The dependent measure was how long toddlers looked at the screen. Figure \@ref(fig:methodfig) shows the design of the experiment.

```{r methodfig, fig.env="figure", fig.width=4.5 ,fig.cap="Study design, habituation, and test phases of the current study. Curly brackets represent within-participant randomized blocks while arrows represent between-participant randomization of trials."}
methodfig <- readJPEG("figs/negcom.jpg")
grid.raster(methodfig)
```

### Participants & Materials
N toddlers in the age range of 18 to 30 months were tested. The experiment only used three images: a tennis ball, a dog, and a shoe (Figure). Images were selected from a free online repository. Using the MB-CDI data available through Wordbank [@frank2016wordbank], we made sure that half of toddlers in that age range produce the corresponding nouns (*ball*, *dog*, and *shoe*) by 18 months and almost all toddlers produce them by age 24 months. 18 linguistic stimuli were recorded corresponding to the 18 combinations of adverbs and nouns. Recordings are available on the study's online repository.

### Procedure

The study consisted of two phases: habituation and test. In habituation trials, toddlers heard positive sentences of the general form "this is [adverb] a [noun]". Adverbs were randomly selected from the following set: *indeed*, *really*, *just*, *only*. After the adverb was selected, nouns were randomly selected from the following set: *ball*, *dog*, *shoe*. Nouns were randomized within the the adverb blocks, and adverb blocks were in turn randomized as well. In each trial, first an attention getter appeard in the middle of the screen. When toddlers looked at the screen the object appeard and the audio of the sentence was played. The sentence was repeated three times. Trial ended when toddlers looked away for more than 2 seconds. All the sentences during the habituation phase were true of the pictures presented on the screen. In ther words, the adverbs did not alter the truth conditions of the statement. The habituation phase continued until toddler's mean looking time for three consecutive trials was reduced to 50% of the mean looking time of their first three trials (Cohen and Gelber, 1975). After infants reached this criterion or were done with all 12 habituation trials, we started the test phase.

The test phase had two between-participant conditions: positive (control) and negative. The positive condition used the adverb *now* and was similar to the habituation phase in that the adverb was positive and did not alter the truth conditions of the statement. The negative condition used the adverb *not* which altered the truth conditions of the statement. Only in such negative trials, the statement was false with respect to the image on the screen. We predicted that if toddlers are sensitive to the truth conditional contributions of the adverb *not*, they will dishabituate in the negative condition but not the positive (control) condition.

## Results

We used `r cite_r("r-references.bib")` for all our analyses.

## Discussion

# Experiment 2

# Appendix

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
